---
title: "Project II Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setting up the enviornment - Install the packages
```{r, results="hide", message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
library(multcomp)
library(knitr)
library(MASS)
library(tools)
import::from(multcomp, glht, mcp, contrMat)
import::from(broom, tidy)
source("util.R") 
```

# Data wrangling with output file generated by software
```{r, message=FALSE, results='asis'}
map_condition <- read_csv("mapcondition.csv")

# read all experiment .csv files
files = list.files(path="./experiment-data", pattern="*.csv", full.names = TRUE)

# this line can be used to read each file individually
# for (i in 1:length(files)) assign(file_path_sans_ext(files[i]), read.csv(file.path("experiment-data", files[i])))

# combine all experiment data into one table
data <- files %>% 
  lapply(read_csv) %>% 
  bind_rows 

# add IV columns
data[, "EO"] <- NA
data[, "TL"] <- NA
data[, "ES"] <- NA
data[, "SS"] <- NA

# map endIndex to condition
for(i in 1:nrow(data)){
  # starting index of table is one
  index_that_maps <- data$endIndex[i] + 1
  row_to_map <- map_condition[index_that_maps,]
  data[i,]$EO <- row_to_map$`Edge orientation`
  data[i,]$TL <- row_to_map$`Target location`
  data[i,]$ES <- row_to_map$`Edge section`
  data[i,]$SS <- row_to_map$`Side section`
}

kable(head(data, n=5), caption="A look at some rows in the experiment data")
```

# Visualizing the data

# Addressing our first hypothesis
##### H1: It is faster to move to targets on the top edge of the screen rather than the bottom edge.
```{r, message=FALSE}
# to address the first hypothesis, we group by Edge Orientation (Horizontal) and Edge Section (Top and Bottom)
data_horizontal_edge <- data[data$EO == "Horizontal",]

# Violin plot  
h1_violin_plot <-
  data_horizontal_edge %>%
  ggplot(aes(x = ES, y = MT)) +
  geom_violin() +
  geom_point(color = "lightgray") +
  labs(title = "Violin plots of edge section on movement time", x = "Edge Section", y = "Movement Time")

h1_violin_plot + stat_summary(fun.y=mean, geom="point", size=2, color="red")
```

# Addressing our second hypothesis
##### H2: For left-handed participants it is faster to move to targets on the right edge of the screen rather than the left edge, whereas for the right-handed participants it is faster to move to targets at the left edge of the screen rather than the right edge.
```{r, message=FALSE}
# to address the second hypothesis, we group by Edge Orientation (Vertical) and Side Section (Dominant and Non-dominant)
data_vertical_edge <- data[data$EO == "Vertical",]

# Violin plot  
h2_violin_plot <-
  data_vertical_edge %>%
  ggplot(aes(x = SS, y = MT)) +
  geom_violin() +
  geom_point(color = "lightgray") +
  labs(title = "Violin plots of side section on movement time", x = "Side Section", y = "Movement Time")

h2_violin_plot + stat_summary(fun.y=mean, geom="point", size=2, color="red")
```